---
title: "02_metrics"
author: "matheusmoroti"
format: html
editor: visual
---

### 0. Libraries

```{r}
library(tidyverse)  # data science handling
library(data.table) # data science handling
library(terra)     # geospatial tool
library(sf) # spatial data package
library(subniche) # subniche
```

### 1. Load data

##### 1.1. Distribution data

```{r}
# Community data from TetrapodTraits
vert_assemblages <- data.table::fread("02_metrics/Tetrapod_360.csv",
                                      stringsAsFactors=T)

# Shapefile
grid_cells_sf <- sf::st_read(dsn = file.path(
    "..", "wallacean_time", "Shapefiles", "TetrapodTraits_community"),
    layer='gridcells_110km'
)

grid_marine <- sf::st_read(dsn = file.path(
    "shapefile"),
    layer='grid_marine_bioclimate'
)
```

##### 1.2. Climate data

```{r}
climate_assemblages <- data.table::fread("02_metrics/SpatialData_110km.csv",
                                      stringsAsFactors=T) 
```

```{r}
# Climate data from CHELSA
# annual mean temperature
Bio01 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio1_1981-2010_V.2.1.tif") 
# temperature seasonality
Bio04 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio4_1981-2010_V.2.1.tif") 
# maximum temperature of the warmest month
Bio05 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio5_1981-2010_V.2.1.tif") 
# minimum temperature of the coldest month
Bio06 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio6_1981-2010_V.2.1.tif") 
# annual precipitation
Bio12 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio12_1981-2010_V.2.1.tif") 
# precipitation seasonality
Bio15 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio15_1981-2010_V.2.1.tif") 
# precipitation of the wettest quarter
Bio16 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio16_1981-2010_V.2.1.tif") 
# precipitation of the driest quarter
Bio17 <- rast("https://os.zhdk.cloud.switch.ch/chelsav2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio17_1981-2010_V.2.1.tif") 
```

##### 1.3. Species list

Herein, we use only species with body size or body mass data observed (non-imputed). For this reason, we use the BAMM trait file for each group. Those files were generated after run *01_extraction.qmd*.

```{r}
amphibia_trait <- read.table(file.path("BAMM",
                                      "Amphibia",
                                      "Amphibia_body_traits.txt"))
reptilia_trait <- read.table(file.path("BAMM",
                                      "Reptilia",
                                      "Reptilia_body_traits.txt"))
aves_trait <- read.table(file.path("BAMM",
                                      "Aves",
                                      "Aves_body_traits.txt"))
mammalia_trait <- read.table(file.path("BAMM",
                                      "Mammalia",
                                      "Mammalia_body_traits.txt"))

all_species <- bind_rows(amphibia_trait,
         reptilia_trait,
         aves_trait,
         mammalia_trait) %>%
 rename(Scientific.Name = V1,
       Body = V2)
```

### 2. Data processing

##### 2.1. Obtain assemblage community

```{r}
vert_assemblages_edit <- vert_assemblages %>%
 mutate(Scientific.Name = str_replace_all(Scientific.Name, "_", " "))

vert_assemblages_matrix <- dcast(
  data = vert_assemblages_edit,
  formula = Cell_Id110 ~ Scientific.Name,
  fun.aggregate = function(x) as.integer(length(x) > 0),  # 1 se presença, 0 se ausência
  value.var = "Cell_Id110",
  fill = 0  # Preenche células sem ocorrência com 0
) 
```

```{r}
# Species list in assemblages data
list <- vert_assemblages_edit %>%
    distinct(Scientific.Name)

# Species list with observed body size/mass
species <- all_species %>% # 31682
 mutate(Scientific.Name = str_replace_all(Scientific.Name, "_", " ")) 

# Species list in both datasets
species_list <- inner_join(list,
                           species,
                           by = "Scientific.Name") # 31671 species
# Species without distribution data
anti_join(species,list, by = "Scientific.Name") 
```

```{r}
cols <- c("Cell_Id110", species_list$Scientific.Name)

vert_assemblages_matrix_filter <- vert_assemblages_matrix %>%
 dplyr::select(all_of(cols)) # 31671 species + cell_id110
```

##### 2.2. Extract climate variables to grids

```{r}
grid_marine_variables <- grid_marine %>%
    select("Cl_I110", "X_meanmean", "X_rangemean", 
           "X_maxmean","X_minmean","geometry") %>%
    st_drop_geometry()

head(vert_assemblages_matrix_filter[1,1])
# Species list in both datasets
sea_communities <- inner_join(vert_assemblages_matrix_filter[,1],
                           grid_marine_variables,
                           by = c("Cell_Id110" = "Cl_I110")) # 31671 
visdat::vis_miss(sea_communities)
nrow(grid_marine_variables)
nrow(sea_communities)

# Species without distribution data
anti_join(vert_assemblages_matrix_filter[,1],
                           grid_marine_variables,
                           by = c("Cell_Id110" = "Cl_I110"))

vert_assemblages_matrix_filter %>%
 filter(Cell_Id110 %in% sea_communities$Cell_Id110)


grids <- vert_assemblages_matrix_filter %>%
 pull("Cell_Id110")
# Only grids with vertebrates
grid_cells_sf_select <- grid_cells_sf %>%
  filter(Cl_I110 %in% grids)
```

Terrestrial habitat

```{r}
valores_bio01 <- extract(Bio01, grid_cells_sf_select, fun = median, na.rm = TRUE)
#valores_bio04 <- extract(Bio04, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio05 <- extract(Bio05, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio06 <- extract(Bio06, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio12 <- extract(Bio12, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio15 <- extract(Bio15, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio16 <- extract(Bio16, grid_cells_sf, fun = median, na.rm = TRUE)
#valores_bio17 <- extract(Bio17, grid_cells_sf, fun = median, na.rm = TRUE)
# rename
names(valores_bio01)[2] <- "Bio01_median"
#names(valores_bio04)[2] <- "Bio04_median"
#names(valores_bio05)[2] <- "Bio05_median"
#names(valores_bio06)[2] <- "Bio06_median"
#names(valores_bio12)[2] <- "Bio12_median"
#names(valores_bio15)[2] <- "Bio15_median"
#names(valores_bio16)[2] <- "Bio16_median"
#names(valores_bio17)[2] <- "Bio17_median"

climate_variables <- bind_cols(valores_bio04, valores_bio05, valores_bio06,
 valores_bio12, valores_bio15, valores_bio16, valores_bio17)
```

Marine habitat

### 3. Niche metrics

##### 3.1. VIF climate variables

```{r}
# Preparing dataset
data_past <- as.data.frame(cbind(bio01 = bio01_past[,3], bio04 = bio04_past[,3], bio07 = bio07_past[,3],
                                 bio11 = bio11_past[,3], bio15 = bio15_past[,3], bio17 = bio17_past[,3]))

# VIF
v1 <- usdm::vifcor(data_past, th=0.5) # identify collinear variables that should be excluded
v1
data_past_select <- usdm::exclude(data_past,v1) # VIF < 3

# Preparing dataset
data_present <- climate_dataset %>% select(BIO01, BIO04, BIO07, BIO11, BIO12, BIO15, BIO17)

v2 <- usdm::vifcor(data_present, th=0.8) # identify collinear variables that should be excluded
v2
data_present_select <- usdm::exclude(data_present,v2) # VIF < 3
```

##### 3.2. Outlying Mean Indexes (OMI)

```{r}
# climate data
#climate_past_trans <- decostand(climate_past, "standardize")
dudi_past <- dudi.pca(climate_past, scale = TRUE, center = TRUE, scan = FALSE)

# niche species
nic_amphibia_past <- niche(dudi_past, amphibia_composition, scann = FALSE)
nic_squamata_past <- niche(dudi_past, squamata_composition, scann = FALSE)
nic_birds_past <- niche(dudi_past, birds_composition, scann = FALSE)
nic_mammals_past <- niche(dudi_past, mammals_composition, scann = FALSE)

# to dataframe
niche_amphibia_traits_past <- as.data.frame(niche.param(nic_amphibia_past))
niche_squamata_traits_past <- as.data.frame(niche.param(nic_squamata_past))
niche_birds_traits_past <- as.data.frame(niche.param(nic_birds_past))
niche_mammals_traits_past <- as.data.frame(niche.param(nic_mammals_past))
```

### 4. Save data

```{r}

# Salvar o OMI e o Tol na pasta traits


# join past and present niche climate for tetrapods
# niche position
climate_niche_amphibia <- inner_join(temp_median_anura, temp_median_anura_present, by="Species")
climate_niche_squamata <- inner_join(temp_median_squamata, temp_median_squamata_present, by="Species")
climate_niche_birds <- inner_join(temp_median_birds, temp_median_birds_present, by="Species")
climate_niche_mammals <- inner_join(temp_median_mammals, temp_median_mammals_present, by="Species")

# save dataset
write.table(climate_niche_amphibia, "climate_data/climate_niche_amphibia.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(climate_niche_squamata, "climate_data/climate_niche_squamata.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(climate_niche_birds, "climate_data/climate_niche_birds.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(climate_niche_mammals, "climate_data/climate_niche_mammals.txt", 
            sep=",", quote=FALSE, row.names = F)


###
# load dataset
# past
climate_niche_range_amphibia_past <- read.table("climate_data/climate_niche_range_amphibia_past.txt",
                                                header=TRUE, sep=",", row.names=NULL)
climate_niche_range_squamata_past <- read.table("climate_data/climate_niche_range_squamata_past.txt",
                                           header=TRUE, sep=",", row.names=NULL)
climate_niche_range_birds_past <- read.table("climate_data/climate_niche_range_birds_past.txt",
                                        header=TRUE, sep=",", row.names=NULL)
climate_niche_range_mammals_past <- read.table("climate_data/climate_niche_range_mammals_past.txt",
                                          header=TRUE, sep=",", row.names=NULL)
# present data
climate_niche_range_amphibia <- read.table("climate_data/climate_niche_range_amphibia.txt",
                                           header=TRUE, sep=",", row.names=NULL)
climate_niche_range_squamata <- read.table("climate_data/climate_niche_range_squamata.txt",
                                           header=TRUE, sep=",", row.names=NULL)
climate_niche_range_birds <- read.table("climate_data/climate_niche_range_birds.txt",
                                        header=TRUE, sep=",", row.names=NULL)
climate_niche_range_mammals <- read.table("climate_data/climate_niche_range_mammals.txt",
                                          header=TRUE, sep=",", row.names=NULL)

# niche range lets add var (max - min)
range_niche_amphibia <- inner_join(climate_niche_range_amphibia_past,
                                   climate_niche_range_amphibia,
                                   by="Species")
range_niche_squamata <- inner_join(climate_niche_range_squamata_past,
                                   climate_niche_range_squamata,
                                   by="Species")
range_niche_birds <- inner_join(climate_niche_range_birds_past,
                                climate_niche_range_birds,
                                by="Species")
range_niche_mammals <- inner_join(climate_niche_range_mammals_past,
                                  climate_niche_range_mammals, 
                                  by="Species")                                          

# save niche breadth
write.table(range_niche_amphibia, "climate_data/niche_range_amphibia_letsaddvar.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(range_niche_squamata, "climate_data/niche_range_squamata_letsaddvar.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(range_niche_birds, "climate_data/niche_range_birds_letsaddvar.txt", 
            sep=",", quote=FALSE, row.names = F)
write.table(range_niche_mammals, "climate_data/niche_range_mammals_letsaddvar.txt", 
            sep=",", quote=FALSE, row.names = F)

# save dataset OMI
# present
write.table(niche_amphibia_traits, "climate_data/niche_omi_amphibia.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_squamata_traits, "climate_data/niche_omi_squamata.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_birds_traits, "climate_data/niche_omi_birds.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_mammals_traits, "climate_data/niche_omi_mammals.txt", 
            sep=",", quote=FALSE, row.names = T)

# past
write.table(niche_amphibia_traits_past, "climate_data/niche_omi_amphibia_past.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_squamata_traits_past, "climate_data/niche_omi_squamata_past.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_birds_traits_past, "climate_data/niche_omi_birds_past.txt", 
            sep=",", quote=FALSE, row.names = T)
write.table(niche_mammals_traits_past, "climate_data/niche_omi_mammals_past.txt", 
            sep=",", quote=FALSE, row.names = T)

```