---
title: "03_analysis"
author: "Matheus Moroti"
format: html
editor: visual
---

### 0. Libraries

```{r}
library(vegan) # normalize dataset
library(picante) # prune sample with phy
library(tidyverse) # data science handling
library(phytools) # phylogenetic data
library(geiger) # phylogenetic plot
library(nlme) # gls analysis
library(cowplot) # plots
library(rr2) # r2 pgls
# PWR analysis
library(pez)
library(phylobase)
library(MCMCglmm)
library(diversitree)
library(ib)
# auxiliar na pwr
library(grid)
library(phylopath)
library(ape)
# plot tree
library(ggtree)
library(treeio)
library(tidytree)
library(ggplot2)
library(TDbook)
library(ggtree)
library(ggnewscale)
```

### 1. Load data

#### 1.1. TetrapodTraits

```{r}
TetraData <- data.table::fread("../traits/TetrapodTraits_1.0.0.csv") %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  select(Scientific.Name, Fos, Ter, Aqu, Arb, Aer, 
         Verticality, Nocturnality,
         ImputedHabitat, Genus, Family, Suborder) 
```

#### 1.2. Climate niche

```{r}
load(file.path(
  "..", "02_metrics", "omi_data", "niche_df.RData"
))
```

#### 1.3. Rate trait evolution

```{r}
###################################### Amphibia ################################
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Amphibia",
            "bodyrate_evolution.RData")
)
rate_body_amphibia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_fenotipica = median_rate)

# OMI Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Amphibia",
            "OMIrate_evolution.RData")
)
rate_omi_amphibia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_omi = median_rate)

# Tol Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Amphibia",
            "tolrate_evolution.RData")
)
rate_tol_amphibia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_tol = median_rate)

###################################### Reptilia ################################
# Morphological evolution
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Reptilia",
            "bodyrate_evolution.RData")
)
rate_body_reptilia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_fenotipica = median_rate)

# OMI Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Reptilia",
            "OMIrate_evolution.RData")
)
rate_omi_reptilia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_omi = median_rate)

# Tol Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Reptilia",
            "tolrate_evolution.RData")
)
rate_tol_reptilia <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_tol = median_rate)

###################################### Birds ###################################
# Morphological evolution
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Aves",
            "bodyrate_evolution.RData")
)
rate_body_birds <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_fenotipica = median_rate)
# OMI Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Aves",
            "OMIrate_evolution.RData")
)
rate_omi_birds <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_omi = median_rate)
# Tol Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Aves",
            "tolrate_evolution.RData")
)
rate_tol_birds <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_tol = median_rate)

#################################### Mammals ###################################
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Mammalia",
            "bodyrate_evolution.RData")
)
rate_body_mammals <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_fenotipica = median_rate)

# OMI Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Mammalia",
            "OMIrate_evolution.RData")
)
rate_omi_mammals <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_omi = median_rate)
# Tol Rate
load(
  file.path("G:", "Meu Drive", "Artigos", "vertebrates_evolution", "Mammalia",
            "tolrate_evolution.RData")
)
rate_tol_mammals <- Reduce(
  function(x, y) left_join(x, y, by = "specie"),
  lista_dfs
) %>%
  mutate(
    median_rate = apply(select(., starts_with("median_tree_")),
                        1, median, na.rm = TRUE)
  ) %>% select(specie, median_rate) %>%
  rename(taxa_tol = median_rate)
```

#### 1.4. Phylogeny

```{r}
amphibia_phy_full <- read.tree(file.path(
  "G:", "Meu Drive", "Artigos", "vertebrates_evolution",
  "Amphibia", "Amphibia_Tree_1.tre"))

reptilia_phy_full <- read.tree(file.path(
  "G:", "Meu Drive", "Artigos", "vertebrates_evolution",
  "Reptilia", "Reptilia_Tree_1.tre"))

birds_phy_full <- read.tree(file.path(
  "G:", "Meu Drive", "Artigos", "vertebrates_evolution",
  "Aves", "Aves_Tree_1.tre"))

mammals_phy_full <- read.tree(file.path(
  "G:", "Meu Drive", "Artigos", "vertebrates_evolution",
  "Mammalia", "Mammalia_Tree_1.tre"))
```

### 2. Preparing data

#### 2.1. Amphibia

```{r}
niche_amphibia_df <- niche_amphibia_terrestrial_df %>%
  mutate(Scientific.Name = rownames(niche_amphibia_terrestrial_df)) %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  select(Scientific.Name, OMI, Tol) 
```

Merge traits

```{r}
rate_niche_amphibia <- left_join(niche_amphibia_df,
          rate_body_amphibia,
          by = c("Scientific.Name" = "specie"))

df_trait_amphibia <- rate_niche_amphibia %>%
  left_join(TetraData, by = "Scientific.Name") %>%
  left_join(rate_tol_amphibia, by = c("Scientific.Name" = "specie")) %>%
  left_join(rate_omi_amphibia, by = c("Scientific.Name" = "specie")) %>%
  remove_missing()
```

Phylogeny to PGLS

```{r}
amphibia_community <- df_trait_amphibia %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

amphibia_phy <- picante::prune.sample(amphibia_community,
                              amphibia_phy_full)
```

Transform data
```{r}
# Os modelos melhoram ao transformar as variaveis
trait_amphibia_log <- df_trait_amphibia %>%
  mutate(
    taxa_fenotipica = scale(log(taxa_fenotipica)),
    OMI    = scale(log(OMI)),
    Tol    = scale(log(Tol+1)),
    taxa_tol   = scale(log(taxa_tol)),
    taxa_omi = scale(log(taxa_omi)),
    Verticality = scale(Verticality^2))

# Normalizando as preditoras
normalize_variables <- vegan::decostand(trait_amphibia_log %>%
                            select(OMI, Tol), #, taxa_tol
                          "normalize")

# Removendo as nao padronizadas e 
# e normalizando as preditoras
trait_amphibia_normalize <- cbind(trait_amphibia_log %>%
                               select(-OMI, -Tol),
                             normalize_variables)
```

#### 2.2. Squamata

```{r}
niche_reptilia_df <- niche_reptilia_terrestrial_df %>%
  mutate(Scientific.Name = rownames(niche_reptilia_terrestrial_df)) %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  select(Scientific.Name, OMI, Tol)
```

Merge traits

```{r}
rate_niche_reptilia <- left_join(niche_reptilia_df,
          rate_body_reptilia,
          by = c("Scientific.Name" = "specie"))

df_trait_reptilia <- rate_niche_reptilia %>%
  left_join(TetraData, by = "Scientific.Name") %>%
  left_join(rate_omi_reptilia, by = c("Scientific.Name" = "specie")) %>%
  left_join(rate_tol_reptilia, by = c("Scientific.Name" = "specie")) %>%
  remove_missing()
```

Phylogeny to PGLS

```{r}
reptilia_community <- df_trait_reptilia %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

reptilia_phy <- picante::prune.sample(reptilia_community,
                              reptilia_phy_full)
```

Transform data
```{r}
# Os modelos melhoram ao transformar as variaveis
trait_reptilia_log <- df_trait_reptilia %>%
  mutate(
    taxa_fenotipica = log(taxa_fenotipica),
    OMI    = scale(log(OMI)),
    Tol    = scale(log(Tol+1)),
    taxa_tol   = scale(log(taxa_tol)),
    taxa_omi = scale(log(taxa_omi)),
    Verticality = scale(Verticality^2))

# Normalizando as preditoras
normalize_variables <- vegan::decostand(trait_reptilia_log %>%
                            select(OMI, Tol), #, taxa_tol
                          "normalize")
# Removendo as nao padronizadas e 
# e normalizando as preditoras
trait_reptilia_normalize <- cbind(trait_reptilia_log %>%
                               select(-OMI, -Tol),
                             normalize_variables)
```

#### 2.3. Birds

```{r}
niche_birds_df <- niche_birds_terrestrial_df %>%
  mutate(Scientific.Name = rownames(niche_birds_terrestrial_df)) %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  select(Scientific.Name, OMI, Tol)
```

Merge traits

```{r}
rate_niche_birds <- left_join(niche_birds_df,
          rate_body_birds,
          by = c("Scientific.Name" = "specie"))

df_trait_birds <- rate_niche_birds %>%
  left_join(TetraData, by = "Scientific.Name") %>%
  left_join(rate_omi_birds, by = c("Scientific.Name" = "specie")) %>%
  left_join(rate_tol_birds, by = c("Scientific.Name" = "specie")) %>%
  remove_missing()
```

Phylogeny to PGLS

```{r}
birds_community <- df_trait_birds %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

birds_phy <- picante::prune.sample(birds_community,
                              birds_phy_full)
```

Transform data
```{r}
# Os modelos melhoram ao transformar as variaveis
trait_birds_log <- df_trait_birds %>%
  mutate(
    taxa_fenotipica = scale(log(taxa_fenotipica)),
    OMI    = scale(log(OMI)),
    Tol    = scale(log(Tol+1)),
    taxa_tol   = scale(log(taxa_tol)),
    taxa_omi = scale(log(taxa_omi)),
    Verticality = scale(Verticality^2))

# Normalizando as preditoras
normalize_variables <- vegan::decostand(trait_birds_log %>%
                            select(OMI, Tol), #, taxa_tol
                          "normalize")
# Removendo as nao padronizadas e 
# e normalizando as preditoras
trait_birds_normalize <- cbind(trait_birds_log %>%
                               select(-OMI, -Tol),
                             normalize_variables)
```

#### 2.4. Mammals

```{r}
niche_mammals_df <- niche_mammals_terrestrial_df %>%
  mutate(Scientific.Name = rownames(niche_mammals_terrestrial_df)) %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  select(Scientific.Name, OMI, Tol)
```

Merge traits

```{r}
rate_niche_mammals <- left_join(niche_mammals_df,
          rate_body_mammals,
          by = c("Scientific.Name" = "specie"))

df_trait_mammals <- rate_niche_mammals %>%
  left_join(TetraData, by = "Scientific.Name") %>%
  left_join(rate_omi_mammals, by = c("Scientific.Name" = "specie")) %>%
  left_join(rate_tol_mammals, by = c("Scientific.Name" = "specie")) %>%
  remove_missing()
```

Phylogeny to PGLS

```{r}
mammals_community <- df_trait_mammals %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

mammals_phy <- picante::prune.sample(mammals_community,
                              mammals_phy_full)
```

Transform data
```{r}
# Os modelos melhoram ao transformar as variaveis
trait_mammals_log <- df_trait_mammals %>%
  mutate(
    taxa_fenotipica = scale(log(taxa_fenotipica)),
    OMI    = scale(log(OMI)),
    Tol    = scale(log(Tol+1)),
    taxa_tol   = scale(log(taxa_tol)),
    taxa_omi = scale(log(taxa_omi)),
    Verticality = scale(Verticality^2))

# Normalizando as preditoras
normalize_variables <- vegan::decostand(trait_mammals_log %>%
                            select(OMI, Tol), #, taxa_tol
                          "normalize")
# Removendo as nao padronizadas e 
# e normalizando as preditoras
trait_mammals_normalize <- cbind(trait_mammals_log %>%
                               select(-OMI, -Tol),
                             normalize_variables)
```

### 3. PhyloPath

#### 3.0. Theorical models

```{r}
# adjust in names
rownames(trait_reptilia_log) <- trait_reptilia_log$Scientific.Name
rownames(trait_amphibia_log) <- trait_amphibia_log$Scientific.Name
rownames(trait_birds_log) <- trait_birds_log$Scientific.Name
rownames(trait_mammals_log) <- trait_mammals_log$Scientific.Name

# testar correlacao entre variaveis
usdm::vifstep(trait_amphibia_log[, c("taxa_fenotipica",
                                     "taxa_omi",
                                     "taxa_tol",
                                     "Verticality")], th=10) 
usdm::vifstep(trait_reptilia_log[, c("taxa_fenotipica",
                                     "taxa_omi",
                                     "taxa_tol",
                                     "Verticality")], th=10) 
usdm::vifstep(trait_birds_log[, c("taxa_fenotipica",
                                  "taxa_omi",
                                  "taxa_tol",
                                  "Verticality")], th=10) 
usdm::vifstep(trait_mammals_log[, c("taxa_fenotipica",
                                    "taxa_omi",
                                    "taxa_tol",
                                    "Verticality")], th=10) 
```

```{r}
models_dir <- define_model_set(
  # Modelo nulo
  null = c(),
  # taxa fenotipica como variavel resposta
  A1 = c(
    taxa_fenotipica ~ taxa_omi + taxa_tol,
    taxa_tol        ~ taxa_omi, 
    taxa_omi        ~ Verticality
  ),
  A2 = c(
    taxa_fenotipica ~ taxa_tol + taxa_omi,
    taxa_omi        ~ taxa_tol, 
    taxa_omi        ~ Verticality
  ),
  A3 = c(
    taxa_fenotipica ~ taxa_omi + taxa_tol,
    taxa_omi        ~ taxa_tol,
    taxa_tol        ~ Verticality
  ),
  A4 = c(
    taxa_fenotipica ~ taxa_omi + taxa_tol,
    taxa_tol        ~ taxa_omi,
    taxa_tol        ~ Verticality
  ),
  A5 = c(
    taxa_fenotipica ~ taxa_omi + taxa_tol + Verticality,
    taxa_tol        ~ taxa_omi,
    taxa_tol        ~ Verticality
  ),
  A6 = c(
    taxa_fenotipica ~ taxa_tol + taxa_omi + Verticality,
    taxa_omi        ~ taxa_tol, 
    taxa_omi        ~ Verticality
  ),
  A7 = c(
    taxa_fenotipica ~ taxa_tol + taxa_omi + Verticality,
    Verticality     ~ taxa_tol,
    taxa_tol ~ taxa_omi),
  A8 = c(
    taxa_fenotipica ~ taxa_tol + taxa_omi + Verticality,
    Verticality     ~ taxa_omi,
    taxa_omi ~ taxa_tol),
  B1 = c(
    taxa_omi        ~ taxa_tol + taxa_fenotipica,
    taxa_fenotipica ~ taxa_tol,
    taxa_tol        ~ Verticality
  ),
  B2 = c(
    taxa_omi        ~ taxa_fenotipica + taxa_tol,
    taxa_tol        ~ taxa_fenotipica,
    taxa_tol        ~ Verticality
  ),
  B3 = c(
    taxa_omi        ~ taxa_tol + taxa_fenotipica,
    taxa_tol        ~ taxa_fenotipica,
    taxa_fenotipica ~ Verticality
  ),
  B4 = c(
    taxa_omi        ~ taxa_tol + taxa_fenotipica,
    taxa_fenotipica ~ taxa_tol,
    taxa_fenotipica ~ Verticality
  ),
  B5 = c(
    taxa_omi        ~ taxa_tol + taxa_fenotipica + Verticality,
    taxa_fenotipica ~ taxa_tol,
    taxa_fenotipica ~ Verticality
  ),
  B6 = c(
    taxa_omi        ~ taxa_fenotipica + taxa_tol + Verticality,
    taxa_tol        ~ taxa_fenotipica,
    taxa_tol        ~ Verticality
  ),
  B7 = c(
    taxa_omi        ~ taxa_fenotipica + taxa_tol + Verticality,
    Verticality     ~ taxa_fenotipica,
    taxa_fenotipica ~ taxa_tol
  ),
  B8 = c(
    taxa_omi        ~ taxa_fenotipica + taxa_tol + Verticality,
    Verticality     ~ taxa_tol,
    taxa_tol        ~ taxa_fenotipica
  ),
  C1 = c(
    taxa_tol        ~ taxa_fenotipica + taxa_omi,
    taxa_omi        ~ taxa_fenotipica,
    taxa_fenotipica ~ Verticality
  ),
  C2 = c(
    taxa_tol        ~ taxa_omi + taxa_fenotipica,
    taxa_fenotipica ~ taxa_omi,
    taxa_fenotipica ~ Verticality
  ),
  C3 = c(
    taxa_tol        ~ taxa_fenotipica + taxa_omi,
    taxa_fenotipica ~ taxa_omi,
    taxa_omi        ~ Verticality
  ),
  C4 = c(
    taxa_tol        ~ taxa_fenotipica + taxa_omi,
    taxa_omi        ~ taxa_fenotipica,
    taxa_omi        ~ Verticality
  ),
  C5 = c(
    taxa_tol        ~ taxa_fenotipica + taxa_omi + Verticality,
    taxa_omi        ~ taxa_fenotipica,
    taxa_omi        ~ Verticality
  ),
  C6 = c(
    taxa_tol        ~ taxa_omi + taxa_fenotipica + Verticality,
    taxa_fenotipica ~ taxa_omi,
    taxa_fenotipica ~ Verticality
  ),
  C7 = c(
    taxa_tol        ~ taxa_omi + taxa_fenotipica + Verticality,
    Verticality     ~ taxa_omi,
    taxa_omi        ~ taxa_fenotipica
  ),
  C8 = c(
    taxa_tol        ~ taxa_omi + taxa_fenotipica + Verticality,
    Verticality     ~ taxa_fenotipica,
    taxa_fenotipica ~ taxa_omi
  ),
  V1 = c(
      Verticality     ~ taxa_fenotipica + taxa_omi,
      taxa_omi        ~ taxa_fenotipica,
      taxa_fenotipica ~ taxa_tol
    ),
    V2 = c(
      Verticality     ~ taxa_omi + taxa_fenotipica,
      taxa_fenotipica ~ taxa_omi + taxa_tol
    ),
    V3 = c(
      Verticality     ~ taxa_fenotipica + taxa_omi,
      taxa_fenotipica ~ taxa_omi,
      taxa_omi        ~ taxa_tol
    ),
    V4 = c(
      Verticality     ~ taxa_fenotipica + taxa_omi,
      taxa_omi        ~ taxa_fenotipica + taxa_tol
    ),
    V5 = c(
      Verticality     ~ taxa_fenotipica + taxa_omi + taxa_tol,
      taxa_omi        ~ taxa_fenotipica + taxa_tol
    ),
    V6 = c(
      Verticality     ~ taxa_omi + taxa_fenotipica + taxa_tol,
      taxa_fenotipica ~ taxa_omi + taxa_tol
    ),
    V7 = c(
      Verticality     ~ taxa_omi + taxa_fenotipica + taxa_tol,
      taxa_tol        ~ taxa_omi,
      taxa_omi        ~ taxa_fenotipica
    ),
    V8 = c(
      Verticality     ~ taxa_omi + taxa_fenotipica + taxa_tol,
      taxa_tol        ~ taxa_fenotipica,
      taxa_fenotipica ~ taxa_omi
    )
)
```

```{r}
plot_model_set(models_dir, 
               labels = c("taxa_fenotipica" = "Morphology",
                          "taxa_tol" = "Niche \nbreadth",
                          "taxa_omi" = "Niche position",
                          "Verticality" = "Verticality"),
               nrow = 8, text_size = 3)
```

#### 3.1. Amphibia
```{r}
result_amphibia <- phylo_path(models_dir, # change models 
                              data = trait_amphibia_log,
                              tree = amphibia_phy,
                              model = 'lambda')
plot(summary(result_amphibia))

# media dos modelos concorrentes
average_amphibia_model <- average(result_amphibia)
# increase the curvature to avoid overlapping edges
plot(average_amphibia_model,
     text_size = 4,
     edge_width = 0.25,
     curvature = 0.1,
     rotation = 0,
     algorithm = "sugiyama")
```

#### 3.2. Squamata
```{r}
result_reptilia <- phylo_path(models_dir, 
                     data = trait_reptilia_log,
                     tree = reptilia_phy,
                     model = 'lambda')
plot(summary(result_reptilia))

# media dos modelos concorrentes
average_rep_model <- average(result_reptilia)

# increase the curvature to avoid overlapping edges
plot(average_rep_model,
  text_size = 4,
  edge_width = 1.25,
  curvature = 0.1,
  rotation = 0,
  algorith = 'mds') 
```

#### 3.3. Birds
```{r}
result_birds <- phylo_path(models_dir, 
                     data = trait_birds_log,
                     tree = birds_phy,
                     model = 'lambda')
plot(summary(result_birds))

# media dos modelos concorrentes
average_birds_model <- average(result_birds)

# increase the curvature to avoid overlapping edges
plot(average_birds_model,
               labels = c("taxa_fenotipica" = "Morphology",
                          "taxa_tol" = "Niche \nbreadth",
                          "taxa_omi" = "Niche position",
                          "Verticality" = "Verticality"),
  text_size = 5,
  edge_width = 1.25,
  curvature = 0.05,
  rotation = 0,
  algorith = 'mds') 
```

#### 3.4. Mammals
```{r}
result_mammals <- phylo_path(models_dir, 
                     data = trait_mammals_log,
                     tree = mammals_phy,
                     model = 'lambda')
plot(summary(result_mammals))

# media dos modelos concorrentes
average_mammals_model <- average(result_mammals)
# increase the curvature to avoid overlapping edges
plot(average_mammals_model,
  labels = c("taxa_fenotipica" = "Morphology",
             "taxa_tol" = "Niche \nbreadth",
             "taxa_omi" = "Niche position",
             "Verticality" = "Verticality"),
  text_size = 5,
  edge_width = 1.25,
  curvature = 0.05,
  rotation = 0,
  algorith = 'mds') 
```


### 4. PhyloPath (Without ImputedHabitat)
#### 4.1. Amphibia
```{r}
# Adjust data
trait_amphibia_log_obs <- trait_amphibia_log %>%
  filter(ImputedHabitat == 0)

amphibia_community_obs <- trait_amphibia_log_obs %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

amphibia_phy_obs <- picante::prune.sample(amphibia_community_obs,
                              amphibia_phy_full)

# Models
result_amphibia_obs <- phylo_path(models_dir, # change models 
                              data = trait_amphibia_log_obs,
                              tree = amphibia_phy_obs,
                              model = 'lambda')
plot(summary(result_amphibia_obs))

# media dos modelos concorrentes
average_amphibia_model_obs <- average(result_amphibia_obs)
# increase the curvature to avoid overlapping edges
plot(average_amphibia_model_obs,
     text_size = 4,
     edge_width = 0.25,
     curvature = 0.1,
     rotation = 0,
     algorithm = "sugiyama")
```

#### 4.2. Squamata
```{r}
# Adjust data
trait_reptilia_log_obs <- trait_reptilia_log %>%
  filter(ImputedHabitat == 0)

reptilia_community_obs <- trait_reptilia_log_obs %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

reptilia_phy_obs <- picante::prune.sample(reptilia_community_obs,
                              reptilia_phy_full)

# Models
result_reptilia_obs <- phylo_path(models_dir, 
                     data = trait_reptilia_log_obs,
                     tree = reptilia_phy_obs,
                     model = 'lambda')
plot(summary(result_reptilia_obs))

# media dos modelos concorrentes
average_rep_model_obs <- average(result_reptilia_obs)

# increase the curvature to avoid overlapping edges
plot(average_rep_model_obs,
  text_size = 4,
  edge_width = 1.25,
  curvature = 0.1,
  rotation = 0,
  algorith = 'mds') 
```

#### 4.3. Birds
```{r}
# Adjust data
trait_birds_log_obs <- trait_birds_log %>%
  filter(ImputedHabitat == 0)

birds_community_obs <- trait_birds_log_obs %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

birds_phy_obs <- picante::prune.sample(birds_community_obs,
                              birds_phy_full)

# Models
result_birds_obs <- phylo_path(models_dir, 
                     data = trait_birds_log_obs,
                     tree = birds_phy_obs,
                     model = 'lambda')
plot(summary(result_birds_obs))

# media dos modelos concorrentes
average_birds_model_obs <- average(result_birds_obs)

# increase the curvature to avoid overlapping edges
plot(average_birds_model_obs,
               labels = c("taxa_fenotipica" = "Morphology",
                          "taxa_tol" = "Niche \nbreadth",
                          "taxa_omi" = "Niche position",
                          "Verticality" = "Verticality"),
  text_size = 5,
  edge_width = 1.25,
  curvature = 0.05,
  rotation = 0,
  algorith = "sugiyama") 
```

#### 4.4. Mammals
```{r}
# Adjust data
trait_mammals_log_obs <- trait_mammals_log %>%
  filter(ImputedHabitat == 0)

mammals_community_obs <- trait_mammals_log_obs %>%
  select(Scientific.Name, taxa_fenotipica) %>%
  pivot_wider(names_from = Scientific.Name, values_from = taxa_fenotipica)

mammals_phy_obs <- picante::prune.sample(mammals_community_obs,
                              mammals_phy_full)

# Models
result_mammals_obs <- phylo_path(models_dir, 
                     data = trait_mammals_log_obs,
                     tree = mammals_phy_obs,
                     model = 'lambda')
plot(summary(result_mammals_obs))

# media dos modelos concorrentes
average_mammals_model_obs <- average(result_mammals_obs)
# increase the curvature to avoid overlapping edges
plot(average_mammals_model_obs,
  labels = c("taxa_fenotipica" = "Morphology",
             "taxa_tol" = "Niche \nbreadth",
             "taxa_omi" = "Niche position",
             "Verticality" = "Verticality"),
  text_size = 5,
  edge_width = 1.25,
  curvature = 0.05,
  rotation = 0,
  algorith = 'mds') 
```


### 5.0. Sensitivity analysis of data

```{r}
library(sensiPhy)

# Dataset complete
TetraData <- data.table::fread(
  "../traits/TetrapodTraits_1.0.0.csv") %>%
  mutate(Scientific.Name = gsub(" ", "_", Scientific.Name)) %>%
  dplyr::select("Scientific.Name", "Class", "Order", 
         "BodyMass_g", "BodyLength_mm",
         "ImputedMass", "ImputedLength",
         "Verticality", "ImputedHabitat")

amphibia_data <- TetraData %>%
  filter(Class == "Amphibia") %>%
  mutate(BodyLength_mm = if_else(ImputedLength == 1, NA_real_, BodyLength_mm),
         Verticality = if_else(ImputedHabitat == 1, NA_real_, Verticality)) %>%
  dplyr::select(Scientific.Name, BodyLength_mm, Verticality,
                ImputedHabitat) %>%
  column_to_rownames(var = "Scientific.Name") 

squamata_data <- TetraData %>%
  filter(Class == "Reptilia") %>%
  filter(!Order %in% c("Crocodylia", "Testudines")) %>%
  mutate(BodyLength_mm = if_else(ImputedLength == 1, NA_real_, BodyLength_mm),
         Verticality = if_else(ImputedHabitat == 1, NA_real_, Verticality)) %>%
  dplyr::select(Scientific.Name, BodyLength_mm, Verticality,
                ImputedHabitat) %>%
  column_to_rownames(var = "Scientific.Name")

aves_data <- TetraData %>%
  filter(Class == "Aves") %>%
  mutate(BodyMass_g = if_else(ImputedMass == 1, NA_real_, BodyMass_g),
         Verticality = if_else(ImputedHabitat == 1, NA_real_, Verticality)) %>%
  dplyr::select(Scientific.Name, BodyMass_g, Verticality,
                ImputedHabitat) %>%
  column_to_rownames(var = "Scientific.Name")

mammals_data <- TetraData %>%
  filter(Class == "Mammalia") %>%
  mutate(BodyMass_g = if_else(ImputedMass == 1, NA_real_, BodyMass_g),
         Verticality = if_else(ImputedHabitat == 1, NA_real_, Verticality)) %>%
  dplyr::select(Scientific.Name, BodyMass_g, Verticality,
                ImputedHabitat) %>%
  column_to_rownames(var = "Scientific.Name")

# Tree complete
tree_amphibia <- read.tree(
  file.path(
    "..", "BAMM", "Amphibia", "amph_shl_new_Posterior_7238.1000.trees"
    )
  )
tree_amphibia <- drop.tip(tree_amphibia[[1]], "Homo_sapiens")
tree_squamata <- read.tree(
  file.path(
    "..", "BAMM", "Reptilia", "squam_shl_new_Posterior_9755.1000.trees"
    )
  )[[1]]
tree_birds <- read.nexus(
  file.path(
    "..", "BAMM", "Aves", "birds_hackett_allspecies.nex"
    )
  )[[1]]
tree_mammals <- read.nexus(
  file.path(
    "..", "BAMM", "Mammalia", 
    "mammals_birth_death_node-dated-completed-trees.nex"
    )
  )[[1]]

# Sensitivity test
sensiPhy_amphibia <- miss.phylo.d(
  amphibia_data,
  tree_amphibia,
  binvar = BodyLength_mm)

sensiPhy_squamata <- miss.phylo.d(
  squamata_data,
  tree_squamata,
  binvar = BodyLength_mm)

sensiPhy_birds <- miss.phylo.d(
  aves_data,
  tree_birds,
  binvar = BodyMass_g)

sensiPhy_mammals <- miss.phylo.d(
  mammals_data,
  tree_mammals,
  binvar = BodyMass_g)

# Does it exists phylogenetic signal in verticality?
mammals_data_without <- mammals_data %>%
  filter(ImputedHabitat == 0) %>%
  select(Verticality)

tree_mammals_without <- prune.sample(t(mammals_data_without),
                                     tree_mammals)

physignal_mammals <- phylosig(tree_mammals_without,
         mammals_data_without$Verticality,
         method="lambda",
         test=TRUE) 
```


### 6.0. Plots
#### 6.1. Plot PCA Climatic niche

```{r}
# Load pca
load(
  "../02_metrics/omi_data/niche_pca.RData"
)

# Function to plot PCA
plot_pca <- function(pca_obj, title = NULL) {
  # Extrair dados dos indivíduos
  ind <- as.data.frame(pca_obj$li)
  ind$ID <- rownames(ind)
  
  # Extrair dados das variáveis
  var <- as.data.frame(pca_obj$co)
  var$varname <- rownames(var)
  
  # Verificar nomes das colunas (para adaptação a diferentes objetos PCA)
  axis1_name <- ifelse("Comp1" %in% colnames(var), "Comp1", 
                       ifelse("Axis1" %in% colnames(var), "Axis1", colnames(var)[1]))
  axis2_name <- ifelse("Comp2" %in% colnames(var), "Comp2", 
                       ifelse("Axis2" %in% colnames(var), "Axis2", colnames(var)[2]))
  
  # Criar o plot
  p <- ggplot() +
    geom_point(data = ind, aes(x = Axis1, y = Axis2), color = "gray30", size = 2) +
    geom_segment(data = var, 
                 aes(x = 0, y = 0, 
                     xend = !!sym(axis1_name), 
                     yend = !!sym(axis2_name)),
                 arrow = arrow(length = unit(0.2, "cm")), color = "red") +
    geom_text(data = var, 
              aes(x = !!sym(axis1_name), 
                  y = !!sym(axis2_name), 
                  label = varname),
              color = "red", vjust = -0.5) +
    theme_minimal() +
    labs(x = "PCA Axis 1", y = "PCA Axis 2",
         title = title) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5)
  
  return(p)
}

amphibia_pca <- plot_pca(dudi_amphibia_terrestre)
reptilia_pca <- plot_pca(dudi_reptilia_terrestre)
birds_pca <- plot_pca(dudi_birds_terrestre)
mammals_pca <- plot_pca(dudi_mammals_terrestre)

cowplot::plot_grid(amphibia_pca, reptilia_pca,
                   birds_pca, mammals_pca)
```

#### 6.2. Phylogeny plots
```{r}
phy <- birds_phy # define a filogenia
trait <- trait_birds_log # define o dataframe com os dados 
group <- "Aves" # save figure name

# Preparing data
tip_order <- phy$tip.label

trait_log_ordered <- trait[
  match(tip_order, trait$Scientific.Name), ] %>%
  #left_join(TetraData[,c("Scientific.Name","Suborder")], by = "Scientific.Name") %>%
  mutate(Suborder = as.factor(Suborder))

rownames(trait_log_ordered) <- phy$tip.label
rate <- trait_log_ordered %>%
  select(taxa_omi, taxa_tol)

## load `tree_anole` and `df_svl` from 'TDbook'
svl <- as.matrix(trait_log_ordered["taxa_fenotipica"])[,1]
fit <- phytools::fastAnc(phy, svl, vars=TRUE, CI=TRUE)

td <- data.frame(node = nodeid(phy,
                               names(svl)),
                               trait = svl)
nd <- data.frame(node = names(fit$ace),
                 trait = fit$ace)

d <- rbind(td, nd)
d$node <- as.numeric(d$node)
tree <- full_join(phy, d, by = 'node')

# adicionar tip labels
# Get the order of visualization of the tip.labels (not the order of tip.labels):
d = fortify(phy)
d = subset(d, isTip)
tree_tiporder<-with(d, label[order(y, decreasing=T)])

#FamilyData<-trait_log_ordered[tree_tiporder,] # reorder rows to match the phylogeny tree_tiporder

# Get the suborder names that will appear in the outer band:
levels(trait_log_ordered$Suborder) # reorder levels to match tiplabels
trait_log_ordered$Suborder<-factor(
  trait_log_ordered$Suborder,
  levels = unique(trait_log_ordered$Suborder))

trait_log_ordered$HighRankLabel<-NA

for(i in 1:nlevels(trait_log_ordered$Suborder)){
  selected_row<-which(trait_log_ordered$Suborder==levels(trait_log_ordered$Suborder)[i])
  if(length(selected_row)<=2){
    trait_log_ordered$HighRankLabel[selected_row[1]]<-paste(trait_log_ordered$Suborder[selected_row[1]])
  } else {
    midrow<-(median(selected_row))
    trait_log_ordered$HighRankLabel[midrow]<-paste(trait_log_ordered$Suborder[midrow])
  }
}

# Create a categorical variable to represent the colour of the families in a outer band:
trait_log_ordered$ColorTaxa<-trait_log_ordered$Suborder
unique_levels<-unique(trait_log_ordered$ColorTaxa)
trait_log_ordered$ColorTaxa<-factor(trait_log_ordered$ColorTaxa, levels=unique_levels)
is.even <- function(x) x %% 2 == 0
trait_log_ordered$ColorTaxa<-as.character(trait_log_ordered$ColorTaxa)

for(i in 1:nlevels(trait_log_ordered$Suborder)){
  # Detect if the row i had its level already replaced  
  if(length(which(trait_log_ordered$ColorTaxa==unique_levels[i]))>=1){
    rows_id<-which(trait_log_ordered$ColorTaxa==unique_levels[i])
    if(is.even(i)==TRUE){ trait_log_ordered[rows_id, ncol(trait_log_ordered)]<-rep("A", length(rows_id)) }
    else { trait_log_ordered[rows_id, ncol(trait_log_ordered)]<-rep("B", length(rows_id)) }
  }
}
trait_log_ordered$HexColors<-NA

trait_log_ordered[which(trait_log_ordered$ColorTaxa=="A"),]$HexColors<-"gray20"
trait_log_ordered[which(trait_log_ordered$ColorTaxa=="B"),]$HexColors<-"gray80"

trait_log_ordered<-trait_log_ordered[phy$tip.label,]

# Plot tree with trait rate 
tree_plot <- ggtree(tree, layout = "circular", ladderize = FALSE, size = 1.0) +
  geom_tree(aes(color=trait), continuous = 'color', size = 2) +
  scale_color_viridis_c(option = "viridis")  # Escala contínua

# plot taxas heatmap
plot <- gheatmap(tree_plot, 
                          rate, # only rate of climatic niche
                          width= 0.1,
                          colnames = FALSE) +
  scale_fill_viridis_c(name="continuous\nvalue", alpha = 1)
# Pegue apenas as pontas da árvore
tips_only <- plot$data %>% filter(isTip)
# Adicione o HighRankLabel correto para cada ponta
tips_only$HighRankLabel <- trait_log_ordered$HighRankLabel[
  match(tips_only$label, rownames(trait_log_ordered))]

# adiciona nova camada
plot <- plot + new_scale_fill()
  
tree_plot_h <- gheatmap(
    plot,
    trait_log_ordered[, "HexColors", drop = FALSE],
    offset = 13,
    # 40 para anfibios
    # 35 para repteis
    # 13 para aves
    # 28 para mammals
    width = 0.025,
    colnames_angle = 90,
    colnames = FALSE,
    color = NULL
  ) +
    scale_fill_identity() +
  geom_tiplab2(
    data = tips_only,
    aes(label = HighRankLabel),
    size = 18.62,
    align = FALSE,
    offset = 17
    # 53 para anfibios
    # 42 para repteis
    # 17 para aves
    # 40 para mammals
  ) + 
  geom_text_repel(data = tips_only,
                  force_pull   = 0, # do not pull toward data points
    nudge_y      = 0.05,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    max.iter = 1e4, max.time = 1,
                   aes(label=HighRankLabel)) + 
    theme(legend.position = "right")

ggsave(file.path("..", "figures",
                 paste0("FigTree2", group, ".png")), 
       plot=tree_plot_h, width=100, height=100, units="cm", 
       dpi = 600, limitsize = FALSE)
```

#### 6.3. Model plots
```{r}
load(
  "model_niche_amphibia_log.RData"
)
load(
  "model_niche_reptilia_log.RData"
)
load(
  "model_niche_birds_log.RData"
)
load(
  "model_niche_mammals.RData"
)


summary(model_niche_amphibia)
summary(model_niche_reptilia)
summary(model_niche_birds)
summary(model_niche_mammals)

models <- list(Amphibia = model_niche_amphibia,
               Squamates = model_niche_reptilia,
               Birds = model_niche_birds, 
               Mammals = model_niche_mammals)

results <- modelsummary::modelplot(models, coef_omit = "Interc", draw=FALSE)

results$term <- factor(results$term, 
                       levels = c("OMI", "Tol", "Arb", 
                                  "Fos",
                                  "Ter","Aqu",
                                  "Aer"),
                       labels = c("Niche position\n(OMI)" ,"Niche breadth\n(Tol)", "Arboreal", 
                                  "Fossorial",
                                  "Terrestrial","Aquatic", 
                                  "Aerial"))

background_data <- data.frame(
  model = unique(results$model),  # agora a coluna se chama 'model'
  fill = c("#bfbbbc", "#bfbbbc", "#bfbbbc", "#bfbbbc")
)

p <- ggplot(results, aes(x = 1, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_rect(data = background_data, 
            aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = fill),
            inherit.aes = FALSE, show.legend = FALSE, alpha = 0.1) +
  geom_pointrange(shape = 21, size = 0.5, fill = "#bfbbbc", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = NULL, y = "Model Coefficients (CI 95%)") +
  facet_grid(term ~ model, scales = "free_x", switch = "y") +
  coord_flip() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    axis.title = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 12),
    axis.line = element_line(colour = "black"),
    axis.ticks.y.left = element_blank(),
    axis.text.x = element_text(),
    axis.text.y = element_blank(),
    plot.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0, hjust = 1, vjust = 0.5, size = 12),
    legend.background = element_rect(colour = 'white', fill = 'white'),
    legend.key = element_blank(),
    legend.title = element_blank(),
    legend.position = "top") +
  scale_fill_identity() 

ggsave(paste0("../figures/fig4coefplot.tiff"), plot=p,
       width=10, height=6, units="in", dpi = 'print')
```
]
#### 6.4. Boxplot plots
```{r}
# Função para gerar o plot
make_box <- function(data, var, fill_color, ylab, xlab) {
  ggplot(data, aes(x = reorder(Suborder, !!sym(var), median),
                   y = !!sym(var))) +
    geom_boxplot(fill = fill_color, color = "black") +
    theme_minimal(base_size = 14) +
    labs(
      x = xlab,
      y = ylab
    ) +
    theme(
      axis.text.x = element_text(angle = 75, hjust = 1, size = 10),  # tamanho menor
      axis.title.x = element_text(size = 12),                       # título do eixo
      axis.title.y = element_text(size = 12)                        # título do eixo
    )
}

# Amphibia
p1 <- make_box(trait_amphibia_log, "taxa_fenotipica", xlab = "", "#F57220",
               bquote("Morphology " ~ sigma^2))
p2 <- make_box(trait_amphibia_log, "taxa_omi",  xlab = "", "#5C7879",
               bquote("Niche position " ~ sigma^2))
p3 <- make_box(trait_amphibia_log, "taxa_tol", xlab = "Suborder", "#2DAEB3",
               bquote("Niche breadth " ~ sigma^2))
# Squamates
p4 <- make_box(trait_reptilia_log, "taxa_fenotipica", xlab = "", "#F57220",
               bquote("Morphology " ~ sigma^2))
p5 <- make_box(trait_reptilia_log, "taxa_omi",  xlab = "", "#5C7879",
               bquote("Niche position " ~ sigma^2))
p6 <- make_box(trait_reptilia_log, "taxa_tol",xlab = "Suborder", "#2DAEB3",
               bquote("Niche breadth " ~ sigma^2))

plot_grid(p1,p4,
          p2,p5,
          p3,p6,
          ncol = 2,
          nrow = 3)


# Aves
p1 <- make_box(trait_birds_log, "taxa_fenotipica", xlab = "", "#F57220",
               bquote("Morphology " ~ sigma^2))
p2 <- make_box(trait_birds_log, "taxa_omi", xlab = "", "#5C7879",
               bquote("Niche position " ~ sigma^2))
p3 <- make_box(trait_birds_log, "taxa_tol", xlab = "Suborder", "#2DAEB3",
               bquote("Niche breadth " ~ sigma^2))
# Mammals
p4 <- make_box(trait_mammals_log, "taxa_fenotipica", xlab = "", "#F57220",
               bquote("Morphology " ~ sigma^2))
p5 <- make_box(trait_mammals_log, "taxa_omi", xlab = "", "#5C7879",
               bquote("Niche position " ~ sigma^2))
p6 <- make_box(trait_mammals_log, "taxa_tol", xlab = "Suborder", "#2DAEB3",
               bquote("Niche breadth " ~ sigma^2))

plot_grid(p1,p4,
          p2,p5,
          p3,p6,
          ncol = 2,
          nrow = 3)

```