---
title: "03_analysis"
author: "Matheus Moroti"
format: html
editor: visual
---

### 0.0. Libraries

```{r}
library(visdat) # missing data check
library(vegan) # normalize dataset
library(picante) # prune sample with phy
library(tidyverse) # data science handling
library(phytools) # phylogenetic data
library(nlme) # gls analysis
library(cowplot) # plots
library(modelsummary) # modelplots
library(rr2)
```

### 0.2. Load data

#### 0.2.1. List of species

```{r}
amphibia_full_list <- read.table("list_of_species/composition_amphibia.txt", h=T) 
squamata_full_list <- read.table("list_of_species/composition_reptiles.txt", h=T) #Roll et al. (2017);
birds_full_list <- read.table("list_of_species/composition_birds.txt", h=T) #BirdLife (2015); 
mammals_full_list <- read.table("list_of_species/composition_mammals.txt", h=T) #IUCN  
```

#### 0.2.2. Climate niche

```{r}
# present climate OMI
amphibia_full_niche <- read.table("climate_niche/niche_omi_amphibia.txt", sep=",", h=T)[,2:3]
amphibia_full_niche$Species <- row.names(amphibia_full_niche)

hist(amphibia_full_niche$OMI)
hist(amphibia_full_niche$Tol)

squamata_full_niche <- read.table("climate_niche/niche_omi_squamata.txt", sep=",", h=T)[,2:3]
squamata_full_niche$Species <- row.names(squamata_full_niche)

hist(squamata_full_niche$OMI)
hist(squamata_full_niche$Tol)

birds_full_niche <- read.table("climate_niche/niche_omi_birds.txt", sep=",", h=T)[,2:3]
birds_full_niche$Species <- row.names(birds_full_niche)

mammals_full_niche <- read.table("climate_niche/niche_omi_mammals.txt", sep=",", h=T)[,2:3]
mammals_full_niche$Species <- row.names(mammals_full_niche)

#past climate OMI
amphibia_full_niche_past <- read.table("climate_niche/niche_omi_amphibia_past.txt", sep=",", h=T)[,2:3]
amphibia_full_niche_past$Species <- row.names(amphibia_full_niche_past)

squamata_full_niche_past <- read.table("climate_niche/niche_omi_squamata_past.txt",sep=",", h=T)[,2:3]
squamata_full_niche_past$Species <- row.names(squamata_full_niche_past)

birds_full_niche_past <- read.table("climate_niche/niche_omi_birds_past.txt", sep=",", h=T)[,2:3]
birds_full_niche_past$Species <- row.names(birds_full_niche_past)

mammals_full_niche_past <- read.table("climate_niche/niche_omi_mammals_past.txt", sep=",", h=T)[,2:3]
mammals_full_niche_past$Species <- row.names(mammals_full_niche_past)
```

#### 0.2.3. Rate trait evolution

```{r}
amphibia_full_rates <- read.table("trait_evolution/amphibia_rates_fully.txt", h=F) 
squamata_full_rates <- read.table("trait_evolution/squamata_rates_fully.txt", h=F) 
birds_full_rates <- read.table("trait_evolution/birds_rates_fully.txt", h=F) 
mammals_full_rates <- read.table("trait_evolution/mammals_rates_fully.txt", h=F) 
```

#### 0.2.4. Phylogeny
```{r}
anura_phy_full <- read.tree("phylogeny/amphibia_fully.tre")
squamata_phy_full <- read.tree("phylogeny/squamata_fully.tre")
birds_phy_full <- read.tree("phylogeny/birds_fully.tre")
mammals_phy_full <- read.tree("phylogeny/mammals_fully.tre")
```

### 1.0. Select pool species of Atlantic forest

#### 1.1. Amphibia
Merge with climate niche (Present data)
```{r}
# present data
# Transform column names to rows
amphibia_list <- amphibia_full_list %>% 
  gather(key = "Species", value = "column_value", Rhinella_azarai:Scinax_cretatus) %>%
  distinct(Species) %>%
  arrange(Species)

# adjustments in nomenclature
amphibia_list$Species <- str_replace_all(amphibia_list$Species, "Scinax_v-signatus", "Scinax_v.signatus") 
amphibia_list$Species <- str_replace_all(amphibia_list$Species, "Scinax_x-signatus", "Scinax_x.signatus")

# join with climate niche
species_climate_niche <- left_join(amphibia_list, amphibia_full_niche , by="Species")

# verifing the species don't merge
anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche)) # 0 NA's
# vis_miss(species_climate_niche) # 0.4 % missing data
#View(species_climate_niche)
```

Merge with climate niche (Past data)
```{r}
# past data
# join with climate niche
species_climate_niche_past <- left_join(amphibia_list, amphibia_full_niche_past , by="Species")

# verifing the species don't merge
# anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_past)) # 0 NA's
# vis_miss(species_climate_niche_past) # 0.4 % missing data
#View(species_climate_niche_past)
```

Merge with trait evolution rate
```{r}
# Merge with trait evolution rate
amphibia_full_rates <- amphibia_full_rates %>% rename(Species = V1, trait_rate = V2)
#View(amphibia_full_rates)

# adjustments in nomenclature
amphibia_full_rates$Species <- str_replace_all(amphibia_full_rates$Species, "Hylomantis_aspera", "Agalychnis_aspera") %>%
  str_replace_all("Hylomantis_granulosa", "Agalychnis_granulosa") %>%
  str_replace_all("Eupemphix_nattereri", "Physalaemus_nattereri") %>%
  str_replace_all("Rana_palmipes", "Lithobates_palmipes") %>%
  str_replace_all("Phyllomedusa_azurea", "Pithecopus_azureus") %>%
  str_replace_all("Phyllomedusa_ayeaye", "Pithecopus_ayeaye") %>%
  str_replace_all("Phyllomedusa_nordestina", "Pithecopus_nordestinus") %>%
  str_replace_all("Phyllomedusa_rohdei", "Pithecopus_rohdei") %>%
  str_replace_all("Phyllomedusa_megacephala", "Pithecopus_megacephalus") %>%
  str_replace_all("Phyllomedusa_hypochondrialis", "Pithecopus_hypochondrialis") %>%
  str_replace_all("Phyllomedusa_rohdei", "Pithecopus_rohdei") %>%
  str_replace_all("Scinax_tripui","Ololygon_tripui") %>% 
  str_replace_all("Scinax_cosenzai", "Ololygon_cosenzai") %>% 
  str_replace_all("Scinax_strigilatus", "Ololygon_strigilata") %>% 
  str_replace_all("Bokermannohyla_claresignata", "Boana_claresignata") %>%
  str_replace_all("Hypsiboas_bandeirantes", "Boana_bandeirantes") %>%
  str_replace_all("Hypsiboas_poaju","Boana_poaju")

# merge
# all species (with NA's)
# amphibia_dataset <- left_join(species_climate_niche, amphibia_full_rates, by="Species")

# only species with phylogeny
amphibia_dataset <- inner_join(species_climate_niche, amphibia_full_rates, by="Species")
amphibia_dataset_past <- inner_join(species_climate_niche_past, amphibia_full_rates, by="Species")

# sum(is.na(amphibia_dataset)) # 57 spp ausentes na filogenia
View(amphibia_dataset)
max(amphibia_dataset$Tol)

# conferind NA's
#View(anti_join(amphibia_dataset, amphibia_full_rates, by="Species")) 
#View(anti_join(amphibia_list, amphibia_full_rates, by="Species")) 
```

Phylogeny to PGLS
```{r}
# species in colnames for prune.sample
amp_community_phy <- amphibia_dataset %>%
  spread(Species, as.numeric(trait_rate))

# adjustments in phy file
anura_phy_full$tip.label <- gsub("Hylomantis_aspera", "Agalychnis_aspera", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Hylomantis_granulosa", "Agalychnis_granulosa", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_rohdei", "Pithecopus_rohdei", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_nordestina", "Pithecopus_nordestinus", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_azurea", "Pithecopus_azureus", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_ayeaye", "Pithecopus_ayeaye", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Eupemphix_nattereri", "Physalaemus_nattereri", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Scinax_tripui", "Ololygon_tripui", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Scinax_strigilatus", "Ololygon_strigilata", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Rana_palmipes", "Lithobates_palmipes", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_megacephala", "Pithecopus_megacephalus", fixed=T, anura_phy_full$tip.label)
anura_phy_full$tip.label <- gsub("Phyllomedusa_hypochondrialis", "Pithecopus_hypochondrialis", fixed=T, anura_phy_full$tip.label)

anura_phy_pgls <- prune.sample(amp_community_phy, anura_phy_full)

```

#### 1.2. Squamata
Merge with climate niche (Present data)
```{r}
# present data
# Transform column names to rows
names(squamata_full_list) <- gsub(".", "_", fixed=T, names(squamata_full_list))
squamata_list <- squamata_full_list %>% 
  gather(key = "Species", value = "column_value", Acanthochelys_radiolata:Xenopholis_undulatus) %>%
  distinct(Species) %>%
  arrange(Species)

squamata_full_niche <- squamata_full_niche %>% mutate(Species = gsub("\\.", "_", Species))

# join with climate niche
species_climate_niche_squamata <- left_join(squamata_list, squamata_full_niche, by="Species")

# verifing the species don't merge
anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_squamata)) # 0 NA's
#vis_miss(species_climate_niche) # 0.0 % missing data
nrow(species_climate_niche_squamata)
```

Merge with climate niche (Past data)
```{r}
# past data
# join with climate niche
squamata_full_niche_past <- squamata_full_niche_past %>% mutate(Species = gsub("\\.", "_", Species))
species_climate_niche_past_squamata <- left_join(squamata_list, squamata_full_niche_past , by="Species")

# verifing the species don't merge
# anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_past_squamata)) # 0 NA's
# vis_miss(species_climate_niche_past) # 0.4 % missing data
#View(species_climate_niche_past)
```

Merge with trait evolution rate
```{r}
# Merge with trait evolution rate
squamata_full_rates <- squamata_full_rates %>% rename(Species = V1, trait_rate = V2)
#View(amphibia_full_rates)

# adjustments in nomenclature
#amphibia_full_rates$Species <- str_replace_all(amphibia_full_rates$Species, "Hylomantis_aspera", "Agalychnis_aspera") 

# merge
# all species (with NA's)
# amphibia_dataset <- left_join(species_climate_niche, amphibia_full_rates, by="Species")

# only species with phylogeny
squamata_dataset <- inner_join(species_climate_niche_squamata, squamata_full_rates, by="Species")
squamata_dataset_past <- inner_join(species_climate_niche_past_squamata, squamata_full_rates, by="Species")

anti_join(species_climate_niche_past_squamata, squamata_full_rates, by="Species")
# sairam poucas espécies, maioria crocodilo e tartaruga
# sum(is.na(amphibia_dataset)) # 57 spp ausentes na filogenia
head(squamata_dataset)
hist(squamata_dataset$trait_rate)

# conferind NA's
#View(anti_join(amphibia_dataset, amphibia_full_rates, by="Species")) 
#View(anti_join(amphibia_list, amphibia_full_rates, by="Species")) 
```

Phylogeny to PGLS
```{r}
# species in colnames for prune.sample
squamata_community_phy <- squamata_dataset %>%
  spread(Species, as.numeric(trait_rate))

# adjustments in phy file
#anura_phy_full$tip.label <- gsub("Hylomantis_aspera", "Agalychnis_aspera", fixed=T, anura_phy_full$tip.label)

squamata_phy_pgls <- prune.sample(squamata_community_phy, squamata_phy_full) # 368
nrow(squamata_dataset) # 368 
```


#### 1.3. Birds
Merge with climate niche (Present data)
```{r}
# present data
# Transform column names to rows
names(birds_full_list) <- gsub(".", "_", fixed=T, names(birds_full_list))

View(birds_full_list)

birds_list <- birds_full_list %>% 
  gather(key = "Species", value = "column_value", Amazona_aestiva:Sporophila_hypoxantha) %>%
  distinct(Species) %>%
  arrange(Species)

birds_full_niche <- birds_full_niche %>% mutate(Species = gsub("\\.", "_", Species))

# join with climate niche
species_climate_niche_birds <- left_join(birds_list, birds_full_niche, by="Species")

# verifing the species don't merge
anti_join(birds_list, birds_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_birds)) # 0 NA's
#vis_miss(species_climate_niche) # 0.0 % missing data
nrow(species_climate_niche_birds)
```

Merge with climate niche (Past data)
```{r}
# past data
# join with climate niche
birds_full_niche_past <- birds_full_niche_past %>% mutate(Species = gsub("\\.", "_", Species))
species_climate_niche_past_birds <- left_join(birds_list, birds_full_niche_past , by="Species")

# verifing the species don't merge
# anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_past_birds)) # 0 NA's
# vis_miss(species_climate_niche_past) # 0.4 % missing data
#View(species_climate_niche_past)
```

Merge with trait evolution rate
```{r}
# Merge with trait evolution rate
birds_full_rates <- birds_full_rates %>% rename(Species = V1, trait_rate = V2)
#View(amphibia_full_rates)

# adjustments in nomenclature
#amphibia_full_rates$Species <- str_replace_all(amphibia_full_rates$Species, "Hylomantis_aspera", "Agalychnis_aspera") 

# merge
# all species (with NA's)
# amphibia_dataset <- left_join(species_climate_niche, amphibia_full_rates, by="Species")

# only species with phylogeny
birds_dataset <- inner_join(species_climate_niche_birds, birds_full_rates, by="Species")
birds_dataset_past <- inner_join(species_climate_niche_past_birds, birds_full_rates, by="Species")

anti_join(species_climate_niche_past_birds, birds_full_rates, by="Species")

# sairam poucas espécies, maioria crocodilo e tartaruga
# sum(is.na(amphibia_dataset)) # 57 spp ausentes na filogenia
# head(squamata_dataset)
# hist(squamata_dataset$trait_rate)

# conferind NA's
#View(anti_join(amphibia_dataset, amphibia_full_rates, by="Species")) 
#View(anti_join(amphibia_list, amphibia_full_rates, by="Species")) 
```

Phylogeny to PGLS
```{r}
# species in colnames for prune.sample
birds_community_phy <- birds_dataset %>%
  spread(Species, as.numeric(trait_rate))

# adjustments in phy file
#anura_phy_full$tip.label <- gsub("Hylomantis_aspera", "Agalychnis_aspera", fixed=T, anura_phy_full$tip.label)

birds_phy_pgls <- prune.sample(birds_community_phy, birds_phy_full) # 695
nrow(birds_dataset) # 695 
```

#### 1.4. Mammals
Merge with climate niche (Present data)
```{r}
# present data
# Transform column names to rows
names(mammals_full_list) <- gsub(".", "_", fixed=T, names(mammals_full_list))

#View(mammals_full_list)

mammals_list <- mammals_full_list %>% 
  gather(key = "Species", value = "column_value", Euryoryzomys_lamia:Callithrix_aurita) %>%
  distinct(Species) %>%
  arrange(Species)

mammals_full_niche <- mammals_full_niche %>% mutate(Species = gsub("\\.", "_", Species))

# join with climate niche
species_climate_niche_mammals <- left_join(mammals_list, mammals_full_niche, by="Species")

# verifing the species don't merge
anti_join(mammals_list, mammals_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_mammals)) # 0 NA's
#vis_miss(species_climate_niche) # 0.0 % missing data
nrow(species_climate_niche_mammals)
```

Merge with climate niche (Past data)
```{r}
# past data
# join with climate niche
mammals_full_niche_past <- mammals_full_niche_past %>% mutate(Species = gsub("\\.", "_", Species))
species_climate_niche_past_mammals <- left_join(mammals_list, mammals_full_niche_past , by="Species")

# verifing the species don't merge
# anti_join(amphibia_list, amphibia_full_niche, by="Species") # 0 rows

sum(is.na(species_climate_niche_past_mammals)) # 0 NA's
# vis_miss(species_climate_niche_past) # 0.4 % missing data
#View(species_climate_niche_past)
```

Merge with trait evolution rate
```{r}
# Merge with trait evolution rate
mammals_full_rates <- mammals_full_rates %>% rename(Species = V1, trait_rate = V2)
#View(amphibia_full_rates)

# adjustments in nomenclature
#amphibia_full_rates$Species <- str_replace_all(amphibia_full_rates$Species, "Hylomantis_aspera", "Agalychnis_aspera") 

# merge
# all species (with NA's)
# amphibia_dataset <- left_join(species_climate_niche, amphibia_full_rates, by="Species")

# only species with phylogeny
mammals_dataset <- inner_join(species_climate_niche_mammals, mammals_full_rates, by="Species")
mammals_dataset_past <- inner_join(species_climate_niche_past_mammals, mammals_full_rates, by="Species")

anti_join(species_climate_niche_past_mammals, mammals_full_rates, by="Species")

# sairam poucas espécies, maioria crocodilo e tartaruga
# sum(is.na(amphibia_dataset)) # 57 spp ausentes na filogenia
# head(squamata_dataset)
# hist(squamata_dataset$trait_rate)

# conferind NA's
#View(anti_join(amphibia_dataset, amphibia_full_rates, by="Species")) 
#View(anti_join(amphibia_list, amphibia_full_rates, by="Species")) 
```

Phylogeny to PGLS
```{r}
# species in colnames for prune.sample
mammals_community_phy <- mammals_dataset %>%
  spread(Species, as.numeric(trait_rate))

# adjustments in phy file
#anura_phy_full$tip.label <- gsub("Hylomantis_aspera", "Agalychnis_aspera", fixed=T, anura_phy_full$tip.label)

mammals_phy_pgls <- prune.sample(mammals_community_phy, mammals_phy_full) # 194
nrow(mammals_dataset) # 194 
```


### 2.0. PGLS
#### 2.1. Amphibia
```{r}
# Join  data
amphibia_dataset_past <- amphibia_dataset_past %>% rename(OMI_past = OMI, Tol_past = Tol)
amphibia_dataset_full <- left_join(amphibia_dataset, amphibia_dataset_past[,1:3], by = 'Species')

# dataset normalizado
amphibia_dataset_full_normalizado <- decostand(amphibia_dataset_full[,2:6], "normalize")
amphibia_dataset_full_normalizado <- cbind(Species = amphibia_dataset_full[,1],amphibia_dataset_full_normalizado)

# pgls models
model_anura_svl_normalizado <- gls(trait_rate ~ OMI + Tol + OMI_past + Tol_past , data=amphibia_dataset_full_normalizado,
                       correlation = corPagel(1,anura_phy_pgls, form = ~Species))

model_anura_svl <- gls(trait_rate ~ OMI + Tol + OMI_past + Tol_past , data=amphibia_dataset_full, 
                       correlation = corPagel(1,anura_phy_pgls, form = ~Species))

# results
summary(model_anura_svl_normalizado)
summary(model_anura_svl)
# check models
plot(resid(model_anura_svl_past) ~ fitted(model_anura_svl_past))
qqnorm(resid(model_anura_svl_past_normal))
qqline(resid(model_anura_svl_past_normal))
plot(model_anura_svl)

library(nlme)
library(lmeInfo)

model_anura_svl$dims

g_mlm(model_anura_svl, p_const = c(0,1), r_const = c(0,1,1,1,1), infotype = "expected")
```

#### 2.2. Squamata
```{r}
# Join  data
squamata_dataset_past <- squamata_dataset_past %>% rename(OMI_past = OMI, Tol_past = Tol)
squamata_dataset_full <- left_join(squamata_dataset, squamata_dataset_past[,1:3], by = 'Species')

# dataset normalizado
squamata_dataset_full_normalizado <- decostand(squamata_dataset_full[,2:6], "normalize")
squamata_dataset_full_normalizado <- cbind(Species = squamata_dataset_full[,1],squamata_dataset_full_normalizado)

# pgls models
model_squamata_svl <- gls((trait_rate*1000) ~ OMI + Tol + OMI_past + Tol_past, data=squamata_dataset_full , 
                       correlation = corPagel(1,squamata_phy_pgls, form = ~Species))

model_squamata_svl_normalizado <- gls(trait_rate ~ OMI + Tol + OMI_past + Tol_past, 
                                      data=squamata_dataset_full_normalizado, 
                                      correlation = corPagel(1,squamata_phy_pgls, form = ~Species))

summary(model_squamata_svl)
summary(model_squamata_svl_normalizado)

qqnorm(resid(model_squamata_svl))
qqline(resid(model_squamata_svl))
```

#### 2.3. Birds
```{r}
# Join  data
birds_dataset_past <- birds_dataset_past %>% rename(OMI_past = OMI, Tol_past = Tol)
birds_dataset_full <- left_join(birds_dataset, birds_dataset_past[,1:3], by = 'Species')

# dataset normalizado
birds_dataset_full_normalizado <- decostand(birds_dataset_full[,2:6], "normalize")
birds_dataset_full_normalizado <- cbind(Species = birds_dataset_full[,1],birds_dataset_full_normalizado)

# pgls models
model_birds_svl <- gls(trait_rate*10 ~ OMI + Tol + OMI_past + Tol_past, data=birds_dataset_full, 
                       correlation = corPagel(1,birds_phy_pgls, form = ~Species))

model_birds_svl_normalizado <- gls(trait_rate ~ OMI + Tol + OMI_past + Tol_past, data=birds_dataset_full_normalizado, 
                       correlation = corPagel(1,birds_phy_pgls, form = ~Species))

summary(model_birds_svl)
summary(model_birds_svl_normalizado)

```

#### 2.4. Mammals
```{r}
# Join  data
mammals_dataset_past <- mammals_dataset_past %>% rename(OMI_past = OMI, Tol_past = Tol)
mammals_dataset_full <- left_join(mammals_dataset, mammals_dataset_past[,1:3], by = 'Species')

# dataset normalizado
mammals_dataset_full_normalizado <- decostand(mammals_dataset_full[,2:6], "normalize")
mammals_dataset_full_normalizado <- cbind(Species = mammals_dataset_full[,1],mammals_dataset_full_normalizado)

# pgls models
model_mammals_svl <- gls((trait_rate*1000) ~ OMI + Tol + OMI_past + Tol_past, data=mammals_dataset_full, 
                       correlation = corPagel(1,mammals_phy_pgls, form = ~Species))

model_mammals_svl_normalizado <- gls(trait_rate ~ OMI + Tol + OMI_past + Tol_past, data=mammals_dataset_full_normalizado, 
                       correlation = corPagel(1,mammals_phy_pgls, form = ~Species))

summary(model_mammals_svl)
summary(model_mammals_svl_normalizado)
```


### 3.0. Plots
```{r}
models <- list(Frogs = model_anura_svl,
               Squamates = model_squamata_svl,
               Birds = model_birds_svl, 
               Mammals = model_mammals_svl)
modelplot(models, coef_omit = 'Interc') + facet_grid(~model)

# per specie
cm <- c('Tol_past' = 'Tolerance past',
        'OMI_past' = 'OMI past', 
        'Tol' = 'Tolerance present',
        'OMI' = 'OMI present')

# frogs
b <- list(geom_vline(xintercept = 0, color = 'black'),
          annotate("rect", alpha = .0,
                   xmin = -.1, xmax = .1, 
                   ymin = -Inf, ymax = Inf),
          geom_point(aes(y = term, x = estimate), alpha = .3, 
                     size = 5, color = 'green4'), theme_classic())

modelplot(model_anura_svl, background = b, coef_omit = 'Interc',  coef_map = cm, linetype = 'dotted')

# squamata
c <- list(geom_vline(xintercept = 0, color = 'black'),
          annotate("rect", alpha = .0,
                   xmin = -5.0, xmax = 5, 
                   ymin = -Inf, ymax = Inf),
          geom_point(aes(y = term, x = estimate), alpha = .3, 
                     size = 5, color = 'green'))

modelplot(model_squamata_svl, background = c, coef_omit = 'Interc',  coef_map = cm, linetype = 'dotted')

# birds
d <- list(geom_vline(xintercept = 0, color = 'black'),
          annotate("rect", alpha = .1,
                   xmin = -.0001, xmax = .0001, 
                   ymin = -Inf, ymax = Inf),
          geom_point(aes(y = term, x = estimate), alpha = .3, 
                     size = 5, color = 'burlywood4'), theme_classic())

modelplot(model_birds_svl, background = d, coef_omit = 'Interc',  coef_map = cm, linetype = 'dotted')

# birds
d <- list(geom_vline(xintercept = 0, color = 'black'),
          annotate("rect", alpha = .0,
                   xmin = -20000, xmax = 20000, 
                   ymin = -Inf, ymax = Inf),
          geom_point(aes(y = term, x = estimate), alpha = .5, 
                     size = 5))

modelplot(model_birds_svl, background = d, coef_omit = 'Interc',  coef_map = cm, linetype = 'dotted')

# mammals
theme_set(
    theme_classic(base_size = 15))

a <- list(geom_vline(xintercept = 0, color = 'black'),
          annotate("rect", alpha = .1,
          xmin = -.0001, xmax = .0001, 
          ymin = -Inf, ymax = Inf),
          geom_point(aes(y = term, x = estimate), alpha = .0, 
          size = 10, color = 'brown4'))

modelplot(model_mammals_svl, background = a, coef_omit = 'Interc',  coef_map = cm, linetype = 'dotted')
``` 

```{r}
models <- list(Frogs = model_anura_svl_normalizado,
               Squamates = model_squamata_svl_normalizado,
               Birds = model_birds_svl_normalizado, 
               Mammals = model_mammals_svl_normalizado)

modelplot(models, coef_omit = 'Interc') + facet_grid(~model)
modelplot(model_anura_svl_normalizado, coef_omit = 'Interc')
modelplot(model_mammals_svl_normalizado, coef_omit = 'Interc')
``` 

Phylogenetic plots
```{r}
# substitute values in edge.length with rate trait evolution
for (i in 1:nrow(amphibia_dataset_full)) {
  species <- amphibia_dataset_full$Species[i]
  edge_length <- (amphibia_dataset_full$trait_rate*1000)[i]
  
  anura_phy_pgls$edge.length[anura_phy_pgls$tip.label == species] <- edge_length
}
write.tree(anura_phy_pgls, "frogs_tree.tre")
tree_teste <- full_join(tree, teste, by = 'Species')

###
tree <- anura_phy_pgls
teste <- amphibia_dataset_full
teste$trait_rate <- teste$trait_rate*1000
View(teste)

tree$Valor <- teste$trait_rate[match(tree$tip.label, teste$Species)]

library(geiger)

rownames(teste) <- teste$Species
chk<-name.check(tree,teste)
summary(chk)

scale_insect.pruned<-drop.tip(scale_insect.tre, chk$tree_not_data)

ln.hosts<-setNames(log(teste$trait_rate),rownames(teste))

host.map<-contMap(tree,ln.hosts,plot=FALSE)
host.map<-setMap(host.map,c("forestgreen","firebrick1"))

plot(host.map,lwd=c(1,5),outline=FALSE,ftype="off", type='fan')


plot(setMap(host.map,c("forestgreen","firebrick1")),
     fsize=c(0.5,0.6),leg.txt="Body mass rate evolution",
     dig=2,outline=FALSE,ftype="off", type='fan', lwd=c(1,5))
```
